AXON. A Selection of Greek Historical Inscriptions. A Database for Research and Teaching

### 2016-01-18 source/21/EAGLE2016_Full_Paper_revised.tex
Paper 21 (De Vido et al) edited to the final stage. Paper also updated in the pdfs folder

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{AXON. A Selection of Greek Historical Inscriptions. A Database for Research and Teaching}
 \author[cafo]{Ivan Matijašić\corref{first}}
\author[cafo]{Stefania De Vido}
\author[cafo]{Silvia Palazzo}
\address[cafo]{Dipartimento di Studi Umanistici, Università Ca' Foscari Venezia}
\cortext[first]{Corresponding author. Email: ivan.matijasic@unive.it}
\begin{document}

\maketitle
\begin{abstract}
The AXON Project has developed a database of Greek historical inscriptions, from the birth of the~polis~in the Archaic
Age to 31 BC. Each entry is provided with the object’s description, a complete lemma, Greek text with critical
apparatus, Italian translation and commentary with keywords and indexes, and updated bibliography. New insights for
data-inclusion have been developed. The database supports enlargement and offers a high degree of searchability. Our
aim is to illustrate the structure, the contents and the solutions we have come up with in the development of the AXON
Project. We will also offer some suggestions for teaching and academic research purposes.

@@

\noindent AXON. A Selection of Greek Historical Inscriptions is a project conceived within the Greek Epigraphy Laboratory
(Director, Prof. Claudia Antonetti), and has been brought into existence with the financial support of the University
Ca’ Foscari of Venice (University Project 2013, Project Coordinator, Prof. Stefania De Vido).

Since October 2014 the members of the AXON Project have been developing a database which includes a great variety of
Greek inscriptions of different chronology, typology, and territory of origin. The most recent advances of traditional
epigraphy as well as the scientific acquisitions in the Digital Humanities have been taken into account. The selection
of texts has been made according to a broader notion of ‘historical’ inscription, including not only significant
military, political, and institutional texts, but also those inscriptions which are essential for the social and
cultural understanding of the Greek world.

AXON includes texts from the birth of the Greek polis in the Archaic Age\footnote{ See the Introduction in \citet{nielsen_inventory_2004}; \citet{hansen_polis:_2006}. } to 31 BC, a chronological frame traditionally related to Greek History (though a future
extension of this chronological limit is not excluded). The epigraphic entries have been prearranged in order to allow
a wide and well-structured description of each document. At the same time, a common and coherent lexicon has been
produced, which will permit an easier indexing of significant words and will make future searches much quicker.

@@

\begin{enumerate}
\item Object type
\item Material
\item Object’s dimensions
\item State of preservation
\item Further descriptive elements
\item Date and context of finding
\item Finding site (modern nation, ancient region, ancient and modern name of the city, if known)
\item Actual location (modern nation, city, museum/archaelogical context, inventory number)
\end{enumerate}
The great majority of these categories can be selected from a given number of options from a pull-down menu. Some
categories – such as Object type, Material, or State of preservation – are directly linked with the corresponding
sections in the EAGLE Vocabularies (see \url{http://www.eagle-network.eu/resources/vocabularies/}). Furthermore, a hyperlink
has been created between the AXON-entries and Pleiades website (\url{http://pleiades.stoa.org}): where the finding site is
known, each entry offers the geographic coordinates and a Googlemaps visualisation. This gives the possibility of
rapidly gathering the information for any single ancient location and allows for searches directly from an interactive
map.

@@

\begin{enumerate}
\item Type of Inscription (with link to EAGLE’s Vocabularies:
\url{http://www.eagle-network.eu/resources/vocabularies/typeins/}. The categories are those used by \citet{guarducci_epigrafia_1987})
\item Text’s structure
\item Writing (Execution technique\footnote{ With link to \url{http://www.eagle-network.eu/resources/vocabularies/writing/}.};
different types of epichoric alphabets according to Kirchhoff’s colour-coded map; Local script\footnote{ Following the
categorisation in \citet{jeffery_local_1990}.}; Palaeographic features and letters’ form\footnote{ The letter-form and glyph-form are
based on the symbols of the font Cardo (\url{http://scholarsfonts.net/cardofnt.html}), but many have been developed by the
AXON Team on the examples of letter-form given in Jeffrey 1990 (see also
\url{http://poinikastas.csad.ox.ac.uk/browseGlyphs.shtml}).}; letters’ heights, description and layout of the text field;
Direction of Text)
\item Language (with an option for any dialect’s peculiarities)
\end{enumerate}

\subsubsection{Genetic lemma \& apparatus criticus}

\noindent The text of each inscription is preceded by a hierarchically arranged lemma (the so called genetic lemma, according to
Louis Robert’s definition\footnote{\citet{robert_carie:_1954}.}) and is followed by the apparatus criticus.

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img002.png}
\caption{Example of Abstract in AXON. The different colours allow a selection of words from a drop-down menu. }
\label{fig:2}
\end{figure}

@@

\begin{enumerate}
\item Persons and names (mainly for ‘historical’ characters)
\item Gods and heroes
\item Place names
\item Geographical names
\item Significant words regarding the history of politics and institutions
\item Other relevant keywords
\item Ancient sources
\end{enumerate}

@@

\noindent Finally, an updated bibliography highlights any previous edition for each entry, as well as all the appropriate
secondary sources. SEG abbreviations have been used for epigraphic corpora and other publications (the section
``materiali'' on the website gives access to a list of all abbreviations, useful for students, as well\footnote{
\url{http://virgo.unive.it/venicepigraphy/axon/public/axon/pagine/materiali} (still being processed).}).

@@

\noindent The AXON Project, as an example of a digital edition of inscriptions (see esp. genetic lemma and apparatus) with a high
degree of clarity for contributors and users, is a useful tool for teaching Greek epigraphy as well as ancient history.
Many contributors are university lecturers / professors of Greek Epigraphy, and the scientific committee includes high
school teachers and instructors in classical languages, making AXON especially well-suited for educational purposes and
for use by students: for engaging them, for example, in the composition of entries. The interoperability of the AXON
website and the cross-references to other Digital Humanities projects are essential elements in the development of this
discipline.

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{AXON. A Selection of Greek Historical Inscriptions. A Database for Research and Teaching}
\author[cafo]{Stefania De Vido}
 \author[cafo]{Ivan Matijašić\corref{first}}
\author[cafo]{Silvia Palazzo}
\address[cafo]{Dipartimento di Studi Umanistici, Università Ca' Foscari Venezia}
\cortext[first]{Corresponding author. Email: ivan.matijasic@unive.it}
\begin{document}

\maketitle
\begin{abstract}
The AXON Project has developed a database of Greek historical inscriptions, from the birth of the~\emph{polis}~in the Archaic
Age to 31 BC. Each entry is provided with the object’s description, a complete lemma, Greek text with critical
apparatus, Italian translation and commentary with keywords and indexes, and updated bibliography. New insights for
data-inclusion have been developed. The database supports enlargement and offers a high degree of searchability. Our
aim is to illustrate the structure, the contents and the solutions we have come up with in the development of the AXON
Project. We will also offer some suggestions for teaching and academic research purposes.

@@

\noindent \emph{AXON. A Selection of Greek Historical Inscriptions} is a project conceived within the Greek Epigraphy Laboratory
(Director, Prof. Claudia Antonetti), and has been brought into existence with the financial support of the University
Ca’ Foscari of Venice (University Project 2013, Project Coordinator, Prof. Stefania De Vido).

Since October 2014 the members of the AXON Project have been developing a database which includes a great variety of
Greek inscriptions of different chronology, typology, and territory of origin. The most recent advances of traditional
epigraphy as well as the scientific acquisitions in the Digital Humanities have been taken into account. The selection
of texts has been made according to a broader notion of `historical' inscription, including not only significant
military, political, and institutional texts, but also those inscriptions which are essential for the social and
cultural understanding of the Greek world.

AXON includes texts from the birth of the Greek \emph{polis} in the Archaic Age\footnote{ See the \emph{Introduction} in \citet{nielsen_inventory_2004}; \citet{hansen_polis:_2006}. } to 31 BC, a chronological frame traditionally related to Greek History (though a future
extension of this chronological limit is not excluded). The epigraphic entries have been prearranged in order to allow
a wide and well-structured description of each document. At the same time, a common and coherent lexicon has been
produced, which will permit an easier indexing of significant words and will make future searches much quicker.

@@

\begin{enumerate}
\renewcommand{\theenumi}{\alph{enumi}}
\item Object type
\item Material
\item Object’s dimensions
\item State of preservation
\item Further descriptive elements
\item Date and context of finding
\item Finding site (modern nation, ancient region, ancient and modern name of the city, if known)
\item Actual location (modern nation, city, museum/archaelogical context, inventory number)
\end{enumerate}
The great majority of these categories can be selected from a given number of options from a pull-down menu. Some
categories – such as Object type, Material, or State of preservation – are directly linked with the corresponding
sections in the EAGLE Vocabularies (see \url{http://www.eagle-network.eu/resources/vocabularies/}). Furthermore, a hyperlink
has been created between the AXON-entries and Pleiades website (\url{http://pleiades.stoa.org}): where the finding site is
known, each entry offers the geographic coordinates and a Googlemaps visualisation. This gives the possibility of
rapidly gathering the information for any single ancient location and allows for searches directly from an interactive
map.

@@

\begin{enumerate}
\renewcommand{\theenumi}{\alph{enumi}}
\item Type of Inscription (with link to EAGLE’s Vocabularies:
\url{http://www.eagle-network.eu/resources/vocabularies/typeins/}. The categories are those used by \citet{Guarducci:1967aa})
\item Text’s structure
\item Writing (Execution technique\footnote{ With link to \url{http://www.eagle-network.eu/resources/vocabularies/writing/}.};
different types of epichoric alphabets according to Kirchhoff’s colour-coded map; Local script\footnote{ Following the
categorisation in \citet{jeffery_local_1990}.}; Palaeographic features and letters’ form\footnote{ The letter-form and glyph-form are
based on the symbols of the font Cardo (\url{http://scholarsfonts.net/cardofnt.html}), but many have been developed by the
AXON Team on the examples of letter-form given in \citealp{jeffery_local_1990} (see also
\url{http://poinikastas.csad.ox.ac.uk/browseGlyphs.shtml}).}; letters’ heights, description and layout of the text field;
Direction of Text)
\item Language (with an option for any dialect’s peculiarities)
\end{enumerate}

\subsubsection{Genetic lemma \& \emph{apparatus criticus}}

\noindent The text of each inscription is preceded by a hierarchically arranged lemma (the so called \emph{genetic lemma}, according to
Louis Robert’s definition\footnote{\citet{robert_carie:_1954}.}) and is followed by the apparatus criticus.

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016FullPaperrevised-img002.png}
\caption{Example of Abstract in AXON. The different colours allow a selection of words from a drop-down menu. }
\label{fig:2}
\end{figure}

@@

\begin{enumerate}
\renewcommand{\theenumi}{\alph{enumi}}
\item Persons and names (mainly for `historical' characters)
\item Gods and heroes
\item Place names
\item Geographical names
\item Significant words regarding the history of politics and institutions
\item Other relevant keywords
\item Ancient sources
\end{enumerate}

@@

\noindent Finally, an updated bibliography highlights any previous edition for each entry, as well as all the appropriate
secondary sources. SEG abbreviations have been used for epigraphic \emph{corpora} and other publications (the section
``materiali'' on the website gives access to a list of all abbreviations, useful for students, as well\footnote{
\url{http://virgo.unive.it/venicepigraphy/axon/public/axon/pagine/materiali} (still being processed).}).

@@

\noindent The AXON Project, as an example of a digital edition of inscriptions (see esp. genetic lemma and \emph{apparatus}) with a high
degree of clarity for contributors and users, is a useful tool for teaching Greek epigraphy as well as ancient history.
Many contributors are university lecturers / professors of Greek Epigraphy, and the scientific committee includes high
school teachers and instructors in classical languages, making AXON especially well-suited for educational purposes and
for use by students: for engaging them, for example, in the composition of entries. The interoperability of the AXON
website and the cross-references to other Digital Humanities projects are essential elements in the development of this
discipline.
